Informing forest carbon inventories under the Paris Agreement using the Global Forest Carbon Database (ForC v4.0)

### 2023-05-23 doc/manuscript/ESSDmanuscript.tex
rewording "accounting" and "transfer" (except in display items)
#39 , #51

\runningtitle{IPCC forest C accounting with ForC}

@@

\textbf{Abstract.} Forests are critical for climate change mitigation
and constitute a substantial portion of planned net emissions reductions
under the 2015 Paris Agreement. Yet, the efficacy of greenhouse gas
mitigation planning and reporting is dependent upon the quality of
available emission factors data, including forest carbon (C) stocks and
changes therein. Tens of thousands of relevant forest C estimates have
been published, yet are not readily accesible to the practitioners
compiling national greenhouse gas inventories. Many of these data have,
however, been compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
accounting if made available through the Emission Factor Database (EFDB)
of the International Panel on Climate Change (IPCC). Here, we develop
and document a process for semi-automated submission of data from ForC
into the EFDB, assess the data available and transferred to date, and
provide recommendations for improving forest data collection, analysis,
and reporting to improve accounting of forest-sector greenhouse gas
emissions and removals. We begin by reconciling terminology and mapping
ForC fields into EFDB. This process required some updates to the ForC
database structure, leading to the release of a new version of ForC
(v4.0; described here). As of May 23, 2023, ForC contained
\textasciitilde19323 independent records relevant to EFDB, 1438 of which
have been submitted to date. Among the data in ForC, there is
disproportionate representation of biomass (particularly aboveground)
stocks, with far fewer records for dead organic matter and soil C, and
relatively few or no records for net annual increments or C fluxes into
(gains) or out of (losses) the IPCC-defined C pools. Geographic
representation is also quite uneven, with the highest densities of
relevant records in temperate forests, and with relatively scant
representation of tropical forests in Africa and Asia. This distribution
of records is generally reflected in the subset of records that have
been submitted to EFDB to date. In the future, forest C estimates in
EFDB can be improved through targeted research to fill critical gaps,
reporting of information required by IPCC, and continued submission of
data from scientific publications to the EFDB. Given that climate change
is rapidly impacting the world's forests, timely reporting of recent
estimates will be especially critical to accurate forest C accounting.

@@

The International Panel on Climate Change (IPCC) provides guidance for
national greenhouse gas inventories for reporting to the United Nations
Framework Convention on Climate Change
\citep[UNFCCC,][]{ipcc_2006_2006, ipcc_2019_2019}. Under this guidance,
greenhouse gas inventories include all managed land, including most of
the world's forest land \citep{ogle_delineating_2018}. The IPCC
inventory guidelines include specific instructions for accounting for
greenhouse gas (mainly CO\textsubscript{2}) exchanges between forest
land and the atmosphere \citep{ipcc_agriculture_2006, ipcc_2019_2019}. A
tiered approach to accounting is employed, where the lowest tier (Tier
1) represents the simplest approach and relies on default parameter
values -- for example, forest carbon (C) stocks values by ecozone
\citep{fao_global_2012} and forest age class derived as the average of
published estimates \citep{ipcc_2019_2019, rozendaal_aboveground_2022}.
Tier 1 values have improved over the years as more of the relevant
underlying data has become available
\citep{requenasuarez_estimating_2019, rozendaal_aboveground_2022}, but
there remains room for continuous improvement as the science advances
and datasets grow and become more widely accessible. For example, the
year following the release of the latest IPCC guidelines, a more
thorough analysis of C accumulation in regrowth forests found that
IPCC's Tier 1 default values underestimated C sequestration by 32\% on
average and failed to capture eight-fold variation within ecozones
\citep{cook-patton_mapping_2020}. In addition, it was revealed that C
stocks in mature African tropical montane forests were two-thirds higher
than the IPCC Tier 1 values for these forests
\citep{cuni-sanchez_high_2021}. This rapid evolution of scientific
information on C cycling in forests is valuable for informing climate
change mitigation efforts but requires improved mechanisms for
communicating the latest information from scientific researchers to the
practitioners who need reliable estimates for greenhouse gas mitigation
planning. Moreover, high variability of forest C cycling within ecozones
\citep[e.g.,][]{cook-patton_mapping_2020, cuni-sanchez_high_2021}
implies that it is useful for practitioners to have access to
locally-specific information, when available.

@@

Here, we (1) review IPCC methods and definitions applied to estimate CO2
emissions and removals from forest in the context of typical forest C
estimation methodologies; (2) describe mapping of ForC to IPCC's EFDB;
(3) describe updates to ForC (ForC v4.0), most of which were implemented
to facilitate data transfer to EFDB; (4) summarize the data in ForC
that's relevant to EFDB and records that have been submitted to date;
and (5) provide recommendations for improving data collection, analysis,
database, and accounting.

@@

\begin{figure}
\includegraphics[width=14cm]{figures_tables/C_variable_mapping} \caption{\textbf{Schematic illustrating the carbon pools quantified under IPCC accounting; corresponding ForC variables, and relationships among them.} For each C pool, we show ForC variables corresponding to the stock, stock change (net annual increment), gain (influx), and loss. Most, but not all, EFDB-relevant ForC variables are shown here. Correspondence of ForC variables to IPCC criteria often depends upon measurement protocols (e.g., minimum stem diameter censused). Additional caveats are as follows: (a,b) branch fall and mortality of stems below the minimum stem diameter censused, which are necessary for a full accounting of dead organic matter production but typically assumed negligible for calculations of biomass change, are excluded by common measurement practice (a) or ForC variable definition (b); (c) assumes that leaf production equals leaf fall, or that changes in foliage biomass are negligble; (d,e) belowground components excluded by common measurement practice (d) or ForC variable definition (e); (f) excludes movement of dead wood into litter through breakage or size reduction; (g) measurements often limited to litter horizon (OL) and may exclude larger branches and stems classified as litter and/or the more decomposed layers of the O horizon.}\label{fig:fig_variable_mapping}
\end{figure}

@@

Prior to releasing ForC v4.0, we executed several quality control
measures. First, we implemented a system of continuous integration using
GitHub Actions \citep[\emph{sensu}][]{kim_implementing_2022} to run some
automatic checks any time the master data files are updated, including
outlier tests and checks for completeness and naming consistency of
records across data files. Second, to improve information on geographic
coordinates, we created a field to record coordinate precision (Appendix
A), and flagged and reviewed records with suspected low precision.
Third, to identify erroneous climate data, we compared ForC climate
values to those extracted from WorldClim version 2.1
\citep{vandepol_identifying_2016, bailey_climwin_2016} based on site
coordinates. Records deviating from WorldClim values by more
variable-specific thresholds (\textgreater5°C for mean annual
temperature, \textgreater7.5°C for mean temperatures of the warmest and
coldest months, or \textgreater1 for log(mean annual precipitation in
mm)) were flagged as requiring review prior to use in analysis or
transfer to EFDB.

@@

Manual review of records was the limiting step for data transfer to
EFDB. We prioritized review of (1) records from the Forest Global Earth
Observatory
\citep[ForestGEO,][]{anderson-teixeira_ctfsforestgeo_2015, davies_forestgeo_2021},
(2) studies with confidence intervals recorded in ForC (because
uncertainty estimates are important to the IPCC), (3) original
publications containing large numbers of EFDB-relevant records, and (4)
records from tropical regions. The latter criteria was motivated by the
fact that although tropical forest is the single most important biome
for climate change mitigation \citep{refs}, ground-based data on
tropical forest C cycling tends to be more scarce due to a variety of
challenges \citep{refs, delima_making_2022}, and tropical countries are
more likely to apply Tier 1 methodology that bases forest C budgets on
internationally defined IPCC default values
\citep{romijn_assessing_2015}.

@@

To submit complete, reviewed ForC records into EFDB, we created R
scripts to restructure ForC records and populate EFDB's bulk import form
(``EFDB bulk import.xlsx''). Criteria for data submission were that (1)
records had been checked against the original study and determined to be
complete and correct, and as originally presented, (2) the original
study presented values in tables or text, as opposed to the values
having been digitized from graphs or calculated based on related
variables, and (3) the records had not previously been sent to EFDB.
Once converted into EFDB format, the records were reviewed and then sent
to the IPCC's Technical Support Unit for inclusion in EFDB. Complete
records needed to be reviewed by the EFDB editorial board prior to
posting in the database -- a process that lags behind records transfer
and had not yet been completed for all records sent as of May 23, 2023.

@@

The mapping of ForC fields into EFDB fields is summarized in Appendix B.
For the majority of fields, contents of the field in ForC was
transferred directly into an EFDB field, either as the only contents of
that field or as part of a composite record. For example, ten ForC
fields describing site location, climate, and edaphic properties all
mapped into the EFDB field \emph{Region/Regional conditions} (Appendix
B). In cases where original studies did not present 95\% confidence
intervals (required by IPCC when available) but did present information
required to calculate these (standard error or n and standard
deviation), we calculated the 95\% confidence intervals and populated
the EFDB field with this information (noting the calculation in the EFDB
field \emph{Comments from Data Provider}). For some fields, simple
conditional logic was used to populate EFDB fields based on ForC
records. For example, for stock variables presented in the original
publication in units of dry organic matter mass (as opposed to C),
several greenhouse gasses (CO\textsubscript{2}, CO, CH\textsubscript{4},
NO, NO\textsubscript{2}, N\textsubscript{2}O) were entered in the EFDB
field indicating the greenhouse gases to which the record could be
pertinent (\emph{Gases} field) because these values could be used in
calculations of greenhouse gas emissions from biomass burning
\citep{ipcc_2006_2006}; otherwise, the only pertinent greenhouse gas
would be CO\textsubscript{2}. There were two cases in which more complex
mapping was required: (1) mapping of C cycle variables (section 4.1.1)
and (2) land classification (section 4.1.2).

@@

Given the lack of public information needed to determine whether lands
are classified as managed
\citep{ogle_delineating_2018, deng_comparing_2021}, and because the
IPCC's definition of managed land is more expansive than is commonly
applied in the scientific literature and hence in ForC, we did not
transfer any classification of land management status from ForC to the
EFDB. However, we do provide auxiliary information that should be useful
in making this determination, including geographical location and
notable disturbance events.

@@

\begin{figure}
\includegraphics[width=15cm]{figures_tables/World_Map_of_sites_with_FAO_and_IPCC_data_sent} \caption{\textbf{Map of sites in ForC shaded by number of independent records relevant to (circles) and transferred to (triangles) EFDB.} Symbols are colored according to the number of records at each site. Underlying map shows FAO ecozones, which are coded as follows: Ba-Boreal coniferous forest, Bb-Boreal tundra woodland, BM-Boreal mountain systems, P-Polar, SBSh-Subtropical steppe, SBWh-Subtropical desert, SCf-Subtropical humid forest, SCs-Subtropical dry forest, SM-Subtropical mountain systems, TAr-Tropical rain forest,  TAwa-Tropical moist deciduous forest, TAwb-Tropical dry forest, TBSh-Tropical shrubland, TBWh-Tropical desert, TeBSk-Temperate steppe, TeBWk-Temperate desert, TeDc-Temperate continental forest, TeDo-Temperate oceanic forest, TeM-Temperate mountain systems, TM-Tropical mountain systems.}\label{fig:fig_map}
\end{figure}

\newpage
\begin{figure}
\includegraphics[width=15cm]{figures_tables/Histogram_n_Relevant_and_Transferred_Records} \caption{\textbf{Histograms of number of independent records in ForC relevant to (grey) and transferred to (black) EFDB, organized by (a) dominant vegetation type, (b) FAO ecozone, (c) continent, and (d) stand age.} For dominant vegetation (a), 'Other' includes deciduous needleleaf, mixed broadleaf- needleleaf, non-woody vegetation (e.g., early successional), and incompletely classified or mixed forest types. For FAO ecozones (b), codes are as listed in the caption of Figure 2.}\label{fig:fig_histograms}
\end{figure}

@@

The records transferred from ForC to EFDB were broadly distributed
across Earth's forests (Fig. 2). However, the density of these records
was very unevenly distributed across continents, biomes, and forest
types and was not proportional to the numbers of relevant records in
ForC (Fig. 3). Rather, the largest number of records came from North
America, followed by Asia, South America, and Africa (Fig. 3c), with the
most represented FAO ecozones being boreal coniferous forest, temperate
continental forest, and temperate mountain systems, followed by tropical
rain forests and moist deciduous forests (Fig. 3b). In terms of dominant
vegetation, by far the most records came from needleleaf evergreen
forests, followed by broadleaf deciduous and broadleaf evergreen (Fig.
3b). The largest records came from mature forests (\textgreater100
years), followed by young and intermediate-aged stands (Fig. 3d).

In terms of variables transferred, records were sent for 19 variables
(or closely-related variable groups), including variables from each C
pool (Table 2). The majority (82\%) of records sent were for C stocks,
including 3\% for total biomass, 53\% for aboveground biomass, 4\% for
components of aboveground biomass (wood or foliage), 4\% for root
biomass, 2\% for components of root biomass (coarse or fine roots), 5\%
for dead wood, 3\% for components of dead wood (standing or fallen), 4\%
for litter (entire O horizon or OL layer component), and 4\% for SOM/
SOC. Increment records totaled 9\% of records sent, virtually all for
aboveground biomass (excepting 4 records for delta.O.horizon). The
remaining 9\% of records sent described fluxes, all of which were either
inputs or outputs to the aboveground biomass pool, a subset of which
also described inputs to the dead wood or litter pool (Table 2, Fig. 1).

@@

There is vast potential to expand forest C data in EFDB by completing
the process of reviewing and submitting data that are already in ForC
(Figs. 2-3). So far, only \textasciitilde7\% of the EFDB-relevant data
in ForC have been submitted to EFDB. Although this process requires
manual review of records, the submission of new records to EFDB is
hugely facilitated by the fact that most pertinent information for each
record is already entered in ForC and can be easily prepared for
submission to EFDB using the system developed here. Future efforts to
review studies for transfer should optimize for representation across
geographic regions, forest types, and variables, giving priority to
those from currently under-represented regions and forest types (Figs.
2-3, Table 2), to records from countries relying on existing data for
their greenhouse gas inventories (Tier 1 or 2 methodology), to the
variables most needed by EFDB users, and to the more contemporary
records.

@@

Several EFDB-relevant variables have not been calculated and presented
as frequently as would be possible given existing forest census data and
minimal extra research effort. For example, aboveground woody mortality
(\emph{woody.mortality\_ag}) and aboveground biomass increment
(\emph{delta.agb}) can be calculated from the same census data as
aboveground woody productivity (\emph{ANPP\_woody}), yet the latter has
received far more research attention, and correspondingly has far more
records in ForC \citetext{\citealp[Table
2,][]{anderson-teixeira_carbon_2021}; \citealp[but
see][]{piponiot_distribution_2022}}. Similarly, live coarse root
biomass, total biomass, and changes in both of these pools could in
theory be estimated in parallel with aboveground biomass, with the
greatest barrier being that allometric models for estimating root
biomass are not as reliable or easily available as are those for
aboveground biomass
\citep{chave_improved_2014, rejou-mechain_biomass_2017, gonzalez-akre_allodb_2022}.
However, while equations for estimating root (and thereby total) biomass
require improvement, they do exist for many forest types
\citep[e.g.,][]{brassard_coarse_2011, chojnacky_updated_2014, waring_overlooking_2017}.
In addition, standing dead trees are captured in most forest censuses
and could be used to estimate standing dead wood, although additional
data on breakage would be needed for accurate accounting. We recommend
that, when possible, researchers calculate and report these variables,
following the reporting guidelines specified in section 6.3.

@@

We recommend that, in order to make research valuable to estimate C
stock changes according to methods provided in the IPCC guidelines,
researchers calculate and report results according to IPCC good practice
(Table 3). It is particularly noteworthy that simple decisions on the
presentation of results will determine whether the records meet the
criteria for inclusion in EFDB. Some examples are as follows: (1)
presenting data only in a figure makes them ineligible for inclusion in
EFDB, whereas presentation in a table or supplementary data file allows
inclusion; (2) direct presentation of all relevant variables allows
inclusion, whereas presenting only components of variables of interest
(e.g., parsing litter into fine woody debris, OL, OF, and OH layers) or
requiring simple mathematical operations to obtain a variable of
interest (e.g., \emph{delta.agb} = \emph{ANPP\_woody} -
\emph{woody.mortality.agb}) disqualifies records from inclusion; (3)
matching IPCC-defined thresholds for defining C pools (Table 1), which
may vary by country, can make the data far more relevant for IPCC
accounting (e.g., using a 10 cm cutoff between dead wood and litter,
presenting soil C to a depth of 30 cm). It should also be emphasized
that reporting of 95\% confidence intervals (or other metrics of error),
when applicable, is highly desirable and makes the data more relevant to
IPCC.

@@

Going forward, forest researchers can make their research more useful
for forest C accounting under IPCC guidelines by calculating and
reporting results in ways that are consistent with methodologies
provided in the IPCC guidelines (Tables 1, 3). In addition, substantial
investments in research and researchers focused on ground-based
measurement of forest C will be required to fill knowledge gaps and
thereby increase the accuracy of forest CO\textsubscript{2} inventories
for forest lands under teh Paris Agreement. This challenge is heightened
by the fact that forests are changing rapidly
\citep[e.g.,][]{mcdowell_pervasive_2020}, and data collected a decade or
more in the past may no longer be relevant. This heightens the need for
an efficient system of making forest C data accessible for national
greenhouse gas inventories. We view the system developed here for
submitting ForC data to the IPCC EFDB as one important step towards that
goal.

@@

\codedataavailability{All code and data are openly available. The ForC
database and associated code is available via the ForC repository within
the ForC-db organization on GitHub (https://github.com/forc-db/ForC),
and the version used here (ForC v4.0) archived in Zenodo (DOI: TBD). The
data and code associated with data transfer to EFDB and preparation of
this manuscript are available via the the IPCC-EFDB-integration
repository within the ForC-db organization on GitHub
(https://github.com/forc-db/IPCC-EFDB-integration).} %% use this section when having data sets and software code available

>>>>

\runningtitle{Informing forest CO\textsubscript{2} inventories with
ForC}

@@

\textbf{Abstract.} Forests are critical for climate change mitigation
and constitute a substantial portion of planned net emissions reductions
under the 2015 Paris Agreement. Yet, the efficacy of greenhouse gas
mitigation planning and reporting is dependent upon the quality of
available emission factors data, including forest carbon (C) stocks and
changes therein. Tens of thousands of relevant forest C estimates have
been published, yet are not readily accessible to the practitioners
compiling national greenhouse gas inventories. Many of these data have,
however, been compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
inventories if made available through the Emission Factor Database
(EFDB) of the International Panel on Climate Change (IPCC). Here, we
develop and document a process for semi-automated submission of data
from ForC into the EFDB, assess the data available and submitted to
date, and provide recommendations for improving forest data collection,
analysis, and reporting to improve inventories of forest-sector
greenhouse gas emissions and removals. We begin by reconciling
terminology and mapping ForC fields into EFDB. This process required
some updates to the ForC database structure, leading to the release of a
new version of ForC (v4.0; described here). As of May 23, 2023, ForC
contained \textasciitilde19323 independent records relevant to EFDB,
1438 of which have been submitted to date. Among the data in ForC, there
is disproportionate representation of biomass (particularly aboveground)
stocks, with far fewer records for dead organic matter and soil C, and
relatively few or no records for net annual increments or C fluxes into
(gains) or out of (losses) the IPCC-defined C pools. Geographic
representation is also quite uneven, with the highest densities of
relevant records in temperate forests, and with relatively scant
representation of tropical forests in Africa and Asia. This distribution
of records is generally reflected in the subset of records that have
been submitted to EFDB to date. In the future, forest C estimates in
EFDB can be improved through targeted research to fill critical gaps,
reporting of information required by IPCC, and continued submission of
data from scientific publications to the EFDB. Given that climate change
is rapidly impacting the world's forests, timely reporting of recent
estimates will be especially critical to accurate forest C inventories.

@@

The International Panel on Climate Change (IPCC) provides guidance for
national greenhouse gas inventories for reporting to the United Nations
Framework Convention on Climate Change
\citep[UNFCCC,][]{ipcc_2006_2006, ipcc_2019_2019}. Under this guidance,
greenhouse gas inventories include all managed land, including most of
the world's forest land \citep{ogle_delineating_2018}. The IPCC
inventory guidelines include specific instructions for inventories for
greenhouse gas (mainly CO\textsubscript{2}) exchanges between forest
land and the atmosphere \citep{ipcc_agriculture_2006, ipcc_2019_2019}. A
tiered approach is employed, where the lowest tier (Tier 1) represents
the simplest approach and relies on default parameter values -- for
example, forest carbon (C) stocks values by ecozone
\citep{fao_global_2012} and forest age class derived as the average of
published estimates \citep{ipcc_2019_2019, rozendaal_aboveground_2022}.
Tier 1 values have improved over the years as more of the relevant
underlying data has become available
\citep{requenasuarez_estimating_2019, rozendaal_aboveground_2022}, but
there remains room for continuous improvement as the science advances
and datasets grow and become more widely accessible. For example, the
year following the release of the latest IPCC guidelines, a more
thorough analysis of C accumulation in regrowth forests found that
IPCC's Tier 1 default values underestimated C sequestration by 32\% on
average and failed to capture eight-fold variation within ecozones
\citep{cook-patton_mapping_2020}. In addition, it was revealed that C
stocks in mature African tropical montane forests were two-thirds higher
than the IPCC Tier 1 values for these forests
\citep{cuni-sanchez_high_2021}. This rapid evolution of scientific
information on C cycling in forests is valuable for informing climate
change mitigation efforts but requires improved mechanisms for
communicating the latest information from scientific researchers to the
practitioners who need reliable estimates for greenhouse gas mitigation
planning. Moreover, high variability of forest C cycling within ecozones
\citep[e.g.,][]{cook-patton_mapping_2020, cuni-sanchez_high_2021}
implies that it is useful for practitioners to have access to
locally-specific information, when available.

@@

Here, we (1) review IPCC methods and definitions applied to estimate
CO\textsubscript{2} emissions and removals from forest in the context of
typical forest C estimation methodologies; (2) describe mapping of ForC
to IPCC's EFDB; (3) describe updates to ForC (ForC v4.0), most of which
were implemented to facilitate data submission to EFDB; (4) summarize
the data in ForC that's relevant to EFDB and records that have been
submitted to date; and (5) provide recommendations as to how the
scientific community can better provide useful data for forest C
inventories under the Paris Agreement.

@@

\begin{figure}
\includegraphics[width=14cm]{figures_tables/C_variable_mapping} \caption{\textbf{Schematic illustrating the carbon pools defined under IPCC Guidelines for national greenhouse gas inventories; corresponding ForC variables, and relationships among them.} For each C pool, we show ForC variables corresponding to the stock, stock change (net annual increment), gain (influx), and loss. Most, but not all, EFDB-relevant ForC variables are shown here. Correspondence of ForC variables to IPCC criteria often depends upon measurement protocols (e.g., minimum stem diameter censused). Additional caveats are as follows: (a,b) branch fall and mortality of stems below the minimum stem diameter censused, which are necessary for a full accounting of dead organic matter production but typically assumed negligible for calculations of biomass change, are excluded by common measurement practice (a) or ForC variable definition (b); (c) assumes that leaf production equals leaf fall, or that changes in foliage biomass are negligble; (d,e) belowground components excluded by common measurement practice (d) or ForC variable definition (e); (f) excludes movement of dead wood into litter through breakage or size reduction; (g) measurements often limited to litter horizon (OL) and may exclude larger branches and stems classified as litter and/or the more decomposed layers of the O horizon.}\label{fig:fig_variable_mapping}
\end{figure}

@@

Prior to releasing ForC v4.0, we executed several quality control
measures. First, we implemented a system of continuous integration using
GitHub Actions \citep[\emph{sensu}][]{kim_implementing_2022} to run some
automatic checks any time the master data files are updated, including
outlier tests and checks for completeness and naming consistency of
records across data files. Second, to improve information on geographic
coordinates, we created a field to record coordinate precision (Appendix
A), and flagged and reviewed records with suspected low precision.
Third, to identify erroneous climate data, we compared ForC climate
values to those extracted from WorldClim version 2.1
\citep{vandepol_identifying_2016, bailey_climwin_2016} based on site
coordinates. Records deviating from WorldClim values by more
variable-specific thresholds (\textgreater5°C for mean annual
temperature, \textgreater7.5°C for mean temperatures of the warmest and
coldest months, or \textgreater1 for log(mean annual precipitation in
mm)) were flagged as requiring review prior to use in analysis or
submission to EFDB.

@@

Manual review of records was the limiting step for data submission to
EFDB. We prioritized review of (1) records from the Forest Global Earth
Observatory
\citep[ForestGEO,][]{anderson-teixeira_ctfsforestgeo_2015, davies_forestgeo_2021},
(2) studies with confidence intervals recorded in ForC (because
uncertainty estimates are important to the IPCC), (3) original
publications containing large numbers of EFDB-relevant records, and (4)
records from tropical regions. The latter criteria was motivated by the
fact that although tropical forest is the single most important biome
for climate change mitigation \citep{refs}, ground-based data on
tropical forest C cycling tends to be more scarce due to a variety of
challenges \citep{refs, delima_making_2022}, and tropical countries are
more likely to apply Tier 1 methodology that bases forest C budgets on
internationally defined IPCC default values
\citep{romijn_assessing_2015}.

@@

To submit complete, reviewed ForC records into EFDB, we created R
scripts to restructure ForC records and populate EFDB's bulk import form
(``EFDB bulk import.xlsx''). Criteria for data submission were that (1)
records had been checked against the original study and determined to be
complete and correct, and as originally presented, (2) the original
study presented values in tables or text, as opposed to the values
having been digitized from graphs or calculated based on related
variables, and (3) the records had not previously been submitted to
EFDB. Once converted into EFDB format, the records were reviewed and
then sent to the IPCC's Technical Support Unit for submission to EFDB.
Complete records needed to be reviewed by the EFDB editorial board and
then posted in the database -- a process that lags behind submission of
records and had not yet been completed for all records sent as of May
23, 2023.

@@

The mapping of ForC fields into EFDB fields is summarized in Appendix B.
For the majority of fields, contents of the field in ForC was copied
directly into an EFDB field, either as the only contents of that field
or as part of a composite record. For example, ten ForC fields
describing site location, climate, and edaphic properties all mapped
into the EFDB field \emph{Region/Regional conditions} (Appendix B). In
cases where original studies did not present 95\% confidence intervals
(required by IPCC when available) but did present information required
to calculate these (standard error or n and standard deviation), we
calculated the 95\% confidence intervals and populated the EFDB field
with this information (noting the calculation in the EFDB field
\emph{Comments from Data Provider}). For some fields, simple conditional
logic was used to populate EFDB fields based on ForC records. For
example, for stock variables presented in the original publication in
units of dry organic matter mass (as opposed to C), several greenhouse
gasses (CO\textsubscript{2}, CO, CH\textsubscript{4}, NO,
NO\textsubscript{2}, N\textsubscript{2}O) were entered in the EFDB field
indicating the greenhouse gases to which the record could be pertinent
(\emph{Gases} field) because these values could be used in calculations
of greenhouse gas emissions from biomass burning \citep{ipcc_2006_2006};
otherwise, the only pertinent greenhouse gas would be
CO\textsubscript{2}. There were two cases in which more complex mapping
was required: (1) mapping of C cycle variables (section 4.1.1) and (2)
land classification (section 4.1.2).

@@

Given the lack of public information needed to determine whether lands
are classified as managed
\citep{ogle_delineating_2018, deng_comparing_2021}, and because the
IPCC's definition of managed land is more expansive than is commonly
applied in the scientific literature and hence in ForC, we did not
include any classification of land management status from ForC in the
records submitted to EFDB. However, we did provide auxiliary information
that should be useful in making this determination, including
geographical location and notable disturbance events.

@@

\begin{figure}
\includegraphics[width=15cm]{figures_tables/World_Map_of_sites_with_FAO_and_IPCC_data_sent} \caption{\textbf{Map of sites in ForC shaded by number of independent records relevant to (circles) and submitted to (triangles) EFDB.} Symbols are colored according to the number of records at each site. Underlying map shows FAO ecozones, which are coded as follows: Ba-Boreal coniferous forest, Bb-Boreal tundra woodland, BM-Boreal mountain systems, P-Polar, SBSh-Subtropical steppe, SBWh-Subtropical desert, SCf-Subtropical humid forest, SCs-Subtropical dry forest, SM-Subtropical mountain systems, TAr-Tropical rain forest,  TAwa-Tropical moist deciduous forest, TAwb-Tropical dry forest, TBSh-Tropical shrubland, TBWh-Tropical desert, TeBSk-Temperate steppe, TeBWk-Temperate desert, TeDc-Temperate continental forest, TeDo-Temperate oceanic forest, TeM-Temperate mountain systems, TM-Tropical mountain systems.}\label{fig:fig_map}
\end{figure}

\newpage
\begin{figure}
\includegraphics[width=15cm]{figures_tables/Histogram_n_Relevant_and_Transferred_Records} \caption{\textbf{Histograms of number of independent records in ForC relevant to (grey) and submitted to (black) EFDB, organized by (a) dominant vegetation type, (b) FAO ecozone, (c) continent, and (d) stand age.} For dominant vegetation (a), 'Other' includes deciduous needleleaf, mixed broadleaf- needleleaf, non-woody vegetation (e.g., early successional), and incompletely classified or mixed forest types. For FAO ecozones (b), codes are as listed in the caption of Figure 2.}\label{fig:fig_histograms}
\end{figure}

@@

The ForC records submitted to EFDB were broadly distributed across
Earth's forests (Fig. 2). However, the density of these records was very
unevenly distributed across continents, biomes, and forest types and was
not proportional to the numbers of relevant records in ForC (Fig. 3).
Rather, the largest number of records came from North America, followed
by Asia, South America, and Africa (Fig. 3c), with the most represented
FAO ecozones being boreal coniferous forest, temperate continental
forest, and temperate mountain systems, followed by tropical rain
forests and moist deciduous forests (Fig. 3b). In terms of dominant
vegetation, by far the most records came from needleleaf evergreen
forests, followed by broadleaf deciduous and broadleaf evergreen (Fig.
3b). The largest records came from mature forests (\textgreater100
years), followed by young and intermediate-aged stands (Fig. 3d).

In terms of variables, records were submitted for 19 variables (or
closely-related variable groups), including variables from each C pool
(Table 2). The majority (82\%) of records sent were for C stocks,
including 3\% for total biomass, 53\% for aboveground biomass, 4\% for
components of aboveground biomass (wood or foliage), 4\% for root
biomass, 2\% for components of root biomass (coarse or fine roots), 5\%
for dead wood, 3\% for components of dead wood (standing or fallen), 4\%
for litter (entire O horizon or OL layer component), and 4\% for SOM/
SOC. Increment records totaled 9\% of records sent, virtually all for
aboveground biomass (excepting 4 records for delta.O.horizon). The
remaining 9\% of records sent described fluxes, all of which were either
inputs or outputs to the aboveground biomass pool, a subset of which
also described inputs to the dead wood or litter pool (Table 2, Fig. 1).

@@

There is vast potential to expand forest C data in EFDB by completing
the process of reviewing and submitting data that are already in ForC
(Figs. 2-3). So far, only \textasciitilde7\% of the EFDB-relevant data
in ForC have been submitted to EFDB. Although this process requires
manual review of records, the submission of new records to EFDB is
hugely facilitated by the fact that most pertinent information for each
record is already entered in ForC and can be easily prepared for
submission to EFDB using the system developed here. Future efforts to
review studies for submission should optimize for representation across
geographic regions, forest types, and variables, giving priority to
those from currently under-represented regions and forest types (Figs.
2-3, Table 2), to records from countries relying on existing data for
their greenhouse gas inventories (Tier 1 or 2 methodology), to the
variables most needed by EFDB users, and to the more contemporary
records.

@@

Several EFDB-relevant variables have not been calculated and presented
as frequently as would be possible given existing forest census data and
minimal extra research effort. For example, aboveground woody mortality
(\emph{woody.mortality\_ag}) and aboveground biomass increment
(\emph{delta.agb}) can be calculated from the same census data as
aboveground woody productivity (\emph{ANPP\_woody}), yet the latter has
received far more research attention, and correspondingly has far more
records in ForC \citetext{\citealp[Table
2,][]{anderson-teixeira_carbon_2021}; \citealp[but
see][]{piponiot_distribution_2022}}. Similarly, live coarse root
biomass, total biomass, and changes in both of these pools could in
theory be estimated in parallel with aboveground biomass, with the
greatest barrier being that allometric models for estimating root
biomass are not as reliable or easily available as are those for
aboveground biomass
\citep{chave_improved_2014, rejou-mechain_biomass_2017, gonzalez-akre_allodb_2022}.
However, while equations for estimating root (and thereby total) biomass
require improvement, they do exist for many forest types
\citep[e.g.,][]{brassard_coarse_2011, chojnacky_updated_2014, waring_overlooking_2017}.
In addition, standing dead trees are captured in most forest censuses
and could be used to estimate standing dead wood, although additional
data on breakage would be needed for accurate estimates. We recommend
that, when possible, researchers calculate and report these variables,
following the reporting guidelines specified in section 6.3.

@@

We recommend that, in order to make research valuable to estimate C
stock changes according to methods provided in the IPCC guidelines,
researchers calculate and report results according to IPCC good practice
(Table 3). It is particularly noteworthy that simple decisions on the
presentation of results will determine whether the records meet the
criteria for inclusion in EFDB. Some examples are as follows: (1)
presenting data only in a figure makes them ineligible for inclusion in
EFDB, whereas presentation in a table or supplementary data file allows
inclusion; (2) direct presentation of all relevant variables allows
inclusion, whereas presenting only components of variables of interest
(e.g., parsing litter into fine woody debris, OL, OF, and OH layers) or
requiring simple mathematical operations to obtain a variable of
interest (e.g., \emph{delta.agb} = \emph{ANPP\_woody} -
\emph{woody.mortality.agb}) disqualifies records from inclusion; (3)
matching IPCC-defined thresholds for defining C pools (Table 1), which
may vary by country, can make the data far more relevant estimating
forest C stock changes according to IPCC guidelines (e.g., using a 10 cm
cutoff between dead wood and litter, presenting soil C to a depth of 30
cm). It should also be emphasized that reporting of 95\% confidence
intervals (or other metrics of error), when applicable, is highly
desirable and makes the data more relevant to IPCC.

@@

Going forward, forest researchers can make their research more useful
for forest C inventories under IPCC guidelines by calculating and
reporting results in ways that are consistent with methodologies
provided in the IPCC guidelines (Tables 1, 3). In addition, substantial
investments in research and researchers focused on ground-based
measurement of forest C will be required to fill knowledge gaps and
thereby increase the accuracy of forest CO\textsubscript{2} inventories
for forest lands under teh Paris Agreement. This challenge is heightened
by the fact that forests are changing rapidly
\citep[e.g.,][]{mcdowell_pervasive_2020}, and data collected a decade or
more in the past may no longer be relevant. This heightens the need for
an efficient system of making forest C data accessible for national
greenhouse gas inventories. We view the system developed here for
submitting ForC data to the IPCC EFDB as one important step towards that
goal.

@@

\codedataavailability{All code and data are openly available. The ForC
database and associated code is available via the ForC repository within
the ForC-db organization on GitHub (https://github.com/forc-db/ForC),
and the version used here (ForC v4.0) archived in Zenodo (DOI: TBD). The
data and code associated with data submission to EFDB and preparation of
this manuscript are available via the the IPCC-EFDB-integration
repository within the ForC-db organization on GitHub
(https://github.com/forc-db/IPCC-EFDB-integration).} %% use this section when having data sets and software code available
